---
title: "DSR Chapter 7"
author: "Alice Tivarovsky"
date: "9/2/2020"
output: html_document
---

_Additional reference used: http://www.sthda.com/english/articles/39-regression-model-diagnostics/161-linear-regression-assumptions-and-diagnostics-in-r-essentials/_

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(broom)
```

## 7.4 General Procedure for regression
```{r import}
album1 = read.delim("./data/AlbumSales1.dat", header = TRUE)

```

```{r}
album_sales = lm(sales ~ adverts, data = album1)
album_sales

summary(album_sales)
```

## 7.8 Multiple Regression 

```{r}
album2 = read.delim("./data/Album Sales 2.dat", header = TRUE)

```

Simple model + multivariable model for album sales: 
```{r}
album_sales2 = lm(sales ~ adverts, data = album2)

album_sales3 = lm(sales ~ adverts + airplay + attract, data = album2)

```

Potentially useful shortcut to add predictors to a model you already computed: 
```{r}
album_sales3 = update(album_sales2, .~. + airplay + attract, data = album2)
```

Model summaries: 
```{r}
summary(album_sales2)
summary(album_sales3)

```

R^2 for simple model = 0.3346
R^2 for multivariable model = 0.6647

To get standardized beta coefficients, use `lm.beta` in the `QuantPsyc` package (didn't bother with it here). 

To get confidence intervals around the betas: 
```{r}
confint(album_sales3)
```

### 7.8.4 Comparing Models

Here we compare `album_sale2` to `album_sales3` using a Partial F test (analogous to ANOVA): 
```{r}
anova(album_sales2, album_sales3)

```

F-statistic = 96.447, p-value < 2.2e-16 (highly significant), meaning model with `airplay` and `attract` is much better. 

## Regression Diagnostics

Use `broom` package for regression diagnostics: 
```{r}
album_sales3_diagnostics = augment(album_sales3)

```

Make a plot:
```{r}
album_sales3_diagnostics %>% 
  ggplot(aes(x = adverts, y = sales)) + 
  geom_point() + 
  stat_smooth(method = lm, se = FALSE) +
  geom_segment(aes(xend = adverts, yend = .fitted))

```
Not sure why the residuals aren't plotting correctly. 

Summary of diagnostics: 
```{r}
broom::glance(album_sales3)
```

### 7.9.2 Outliers and Influential cases

Outliers: residuals, standardized residuals, and studentized residuals
```{r}
resid_3 = resid(album_sales3)
std_resid_3 = rstandard(album_sales3)
stud_resid_3 = rstudent(album_sales3)

# combine into a dataframe: 
diag_df = tibble(residual = resid_3, std_residual = std_resid_3, student_resid = stud_resid_3)

```

Plotting diagnostics: 
```{r}
plot(album_sales3)

```

