---
title: "Discovering Stats with R - Chapter 5"
author: "Alice Tivarovsky"
date: "8/9/2020"
output: html_document
---
## Setup
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(car)

```

## 5.5 Assumption of Normality

First we need to fix the outlier from the Download festival dataset. 

```{r}
festival_data = read.delim("DownloadFestival.dat", header = TRUE)

festival_data$day1 = replace(festival_data$day1, festival_data$day1 == 20.02, 2.02)

```

### Comparing to normal curve (histograms)

Plotting histograms for each of the three days of the festival. 

Day 1:

```{r}
hist_day1 = 
  festival_data %>% 
  ggplot(aes(day1)) + geom_histogram(aes(y = ..density..)) + 
  labs(x = "Hygiene score on day 1", y = "Density")

## Adding a normal to day1 curve using stat_function
hist_day1 + stat_function(fun = dnorm, 
                          args = list(mean = mean(festival_data$day1, na.rm = TRUE), 
                                      sd = sd(festival_data$day1, na.rm = TRUE)))

```

Day 2: 

```{r}
hist_day2 = 
  festival_data %>% 
  ggplot(aes(day2)) + geom_histogram(aes(y = ..density..)) + 
  labs(x = "Hygiene score on day 2", y = "Density") +
  stat_function(fun = dnorm, 
                args = list(mean = mean(festival_data$day2, na.rm = TRUE), 
                            sd = sd(festival_data$day2, na.rm = TRUE)))

```

Day 3:

```{r}
hist_day3 = 
  festival_data %>% 
  ggplot(aes(day3)) + geom_histogram(aes(y = ..density..)) + 
  labs(x = "Hygiene score on day 3", y = "Density") +
  stat_function(fun = dnorm, 
                args = list(mean = mean(festival_data$day3, na.rm = TRUE), 
                            sd = sd(festival_data$day3, na.rm = TRUE)))

```

### Q-Q Plots

Q-Q plots compare your data to the same quantiles in a perfect normal curve. 

```{r}
qqplot_day1 = qplot(sample = festival_data$day1)

```

### 5.5.2 Quantifying Normality with numbers

```{r}
# for one vector
psych::describe(festival_data$day1)

# more than one vector
psych::describe(festival_data[, c("day1", "day2", "day3")])

```
### 5.5.3 Looking at groups of data

#### The `by()` function 

Looking at groups within a dataset requires us to use `by()` function. Here we load in a new dataset, `RExam`, which gives exam scores, computer literacy scores, % of lectures attended, and whether the student went to Sussex or another university. We will change the `uni` variable type to factor. 

```{r}
rexam = read.delim("RExam.dat", header = TRUE) %>%
  mutate(uni = as.factor(uni)) %>% 
  mutate(uni = recode(uni, "0" = "Duncetown", "1" = "Sussex"))
  

```

Run histograms for exam, computer, and lectures. 

```{r}
hist_exam = 
  rexam %>% 
  ggplot(aes(exam)) +
  geom_histogram(aes(y =..density..)) + 
  stat_function(fun = dnorm, args = list(mean = mean(rexam$exam, na.rm = TRUE), sd = sd(rexam$exam, na.rm = TRUE)))

hist_comp =
  rexam %>% 
  ggplot(aes(computer)) +
  geom_histogram(aes(y = ..density..)) +
  stat_function(fun = dnorm, args = list(mean = mean(rexam$computer, na.rm = TRUE), sd = sd(rexam$computer, na.rm = TRUE)))

hist_lecture = 
  rexam %>% 
  ggplot(aes(lectures)) +
  geom_histogram(aes(y = ..density..), binwidth = 5) + 
  stat_function(fun = dnorm, args = list(mean = mean(rexam$lectures, na.rm = TRUE), sd = sd(rexam$lectures, na.rm = TRUE)))

hist_exam
hist_comp
hist_lecture

psych::describe(rexam[, c("exam", "computer", "lectures")])

```
Using the `by()` function to split our analysis by those at Sussex and those at Duncetown. 
```{r}
by(data = rexam$exam, INDICES = rexam$uni, FUN = psych::describe)

```

You can also use `by()` with more than one variable. 
```{r}
by(data = rexam[, c("exam", "numeracy")], INDICES = rexam$uni, FUN = psych::describe)

```

#### Subsetting

Another way to analyze groups is to use `subset()`
```{r}
dunce_data = subset(rexam, rexam$uni == "Duncetown")
sussex_data = subset(rexam, rexam$uni == "Sussex")

```

Now we can create separate histograms for the subsets. 
```{r}
hist_dunce_data = 
  dunce_data %>% 
  ggplot(aes(computer)) + 
  geom_histogram(aes(y = ..density..), binwidth = 2) +
  stat_function(fun = dnorm, args = list(mean = mean(rexam$computer, na.rm = TRUE), sd = sd(rexam$computer, na.rm = TRUE)))

hist_sussex_data = 
  sussex_data %>% 
  ggplot(aes(computer)) + 
  geom_histogram(aes(y = ..density..), binwidth = 2) +
  stat_function(fun = dnorm, args = list(mean = mean(rexam$computer, na.rm = TRUE), sd = sd(rexam$computer, na.rm = TRUE)))

hist_dunce_data
hist_sussex_data

```
## 5.6 Testing Normality

### Shapiro Wilk Test
Compares scores from our distribution with those of a normal distribution with the same mean and standard deviation. 

```{r}
shapiro.test(rexam$exam)

```

Can also be done on subsets:
```{r}
by(rexam$exam, INDICES = rexam$uni, FUN = shapiro.test)

```

## 5.7 Testing homogeneity of variance (Levene's Test)

Levene's test = basically ANOVA across all categories. Null hypothesis is that variances across groups are equal --> so p-value < 0.05 means variances across groups are significantly different. 

NOTE: `leveneTest()` function is part of the `car` library.

```{r}
leveneTest(rexam$exam, rexam$uni)
```

## 5.8 Correcting Problems in the Data

### Dealing with outliers
* Remove them
* Transform them
* Change the scores

### `is.na()`
```{r}
festival_data$na_day_2 = is.na(festival_data$day2)

sum(festival_data$na_day_2)

```


### `ifelse()`

* ifelse(condition, what happens if TRUE, what happens if FALSE)
* Good ifelse example with rowMeans and calculation of missing values

```{r}
# first get missingness indicator
festival_data$missing = rowSums(cbind(is.na(festival_data$day1), 
                                      is.na(festival_data$day2),
                                      is.na(festival_data$day3)))

# next calculate means across the three days, excluding missing values
festival_data$meam_hygiene = ifelse(festival_data$missing < 2, NA, 
                                    rowMeans(cbind(festival_data$day1,
                                                   festival_data$day2, 
                                                   festival_data$day3), 
                                             na.rm = TRUE))
festival_data

```

## Tasks

Task 1
```{r}
chick_flick = read.delim("ChickFlick.dat", header = TRUE)

# histogram
chick_flick %>% 
  ggplot(aes(arousal)) +
  geom_histogram(aes(y = ..density..))

#test for normality
shapiro.test(chick_flick$arousal)

# test for homogeneity of variance (Levene)
leveneTest(chick_flick$arousal, group = chick_flick$film, center = median)
```

Task 2
```{r}
rexam$log_numeracy = log(rexam$numeracy)

rexam %>% 
  ggplot(aes(log_numeracy)) + 
  geom_histogram(aes(y = ..density..), binwidth = 0.2)

```

